---
title: ""
editor: source
---

# Exercise 3

All exercises live in the `exercise/` folder. Navigate to that folder in the `Files` pane and click on `03_exercise.qmd`. 

::: {.callout-note}

This document is different than a `.R` script - it is called a Quarto document. Everything outside the {r} code chunks is Markdown language, and everything inside of it is R. For our purposes, all you need to know is to write the code inside of the {r} code chunks and all non-code outside of it (like answers to test questions). 

:::

## Exercise 3.1 

::: panel-tabset
### Problem 

- What is your current directory? 

- Locate the "UADAggs_nat_v3_0.sas7bdat" file within the "lesson_code", "02_lesson", and "data" subdirectories. Read that file into R using a relative file path. 

- Compared to the `.parquet` version of the data, how much larger is the `.sas7bdat` file? (You don't need R for this)

### Solution

- What is your current directory? 

<br>

You can check your current directory with the `getwd()` functino.

<br>

```{r}
#| eval: false

getwd()
```

<br>

- Locate the "UADAggs_nat_v3_0.sas7bdat" file within the "lesson_code", "02_lesson", and "data" subdirectories. Read that file into R using a relative file path. 

<br>

```{r}
#| eval: false

library(haven)
read_sas("lesson_code/02_lesson/data/UADAggs_nat_V3_0.sas7bdat")
```

<br>


- Compared to the `.parquet` version of the data, how much larger is the `.sas7bdat` file? (You don't need R for this)  

The `.parquet` file is 778 KB while the `.sas7bdat` file is 81,892 KB. So the `.sas7bdat` file is more than 105 times larger. 

:::

## Exercise 3.2 - Challenge

::: panel-tabset

### Problem

- Set your current directory to "lesson_code/02_lesson/images". 

- Now read in the same file from Exercise 3.1 using a relative file path. 

- Set your current directory back to the `intro-to-r` folder. 

<br>

### Solution

- Set your current directory to "lesson_code/02_lesson/images". 

<br>
```{r}
#| eval: false

setwd("lesson_code/02_lesson/images")
```
<br>

- Now read in the same file from Exercise 3.1 using a relative file path. 
```{r}
#| eval: false

library(haven)
read_sas("../data/UADAggs_nat_V3_0.sas7bdat")
```
<br>

- Set your current directory back to the `intro-to-r` folder. 

<br>

```{r}
#| eval: false

setwd("../../.../")
```

:::

