---
title: ""
editor: source
---

# R Basics

## Comments

Comments in R are created with the `#` character. We use comments to give the reader of code (which is sometimes yourself in 6 months) an indication of what you are doing and why. Use them frequently to explain what you are doing to those reading your code. 

For multi-line comments, highlight the lines you would like to comment and `Shift` + `CTRL` + `C`. 

<br> 

```{r}
# This is a comment

# This comment
# is multi-line

```

<br> 

## Assignment 

We've already discussed that we can use R like a calculator. 

<br> 

```{r}
# use R like a calculator
3 + 4
```

<br>

The `<-` operator allows you to assign whateer is on the right hand side to an object name on the left hand side in the format of `<obj_name> <- <expr>`. This does two things: 

<br> 

- Creates an object. You can think of an object like you would in real life, but in the virtual world (basically any kind of noun). 
- Creates a name that points to that object. 

<br> 

This allows you to reference the object that the object name points to without needing to type in the values again. 

<br> 

```{r}
#assign values to an object reference
var <- 3 + 4

var
```

<br> 

The `<-` operator is used as opposed to most programming languages where the default assignment is `=`. In R, `=` does work for assignment, but it is best practice to use `<-`. 

<br> 

::: {.callout-note}

If you type th eobject name without assignment, it wil print the result automatically. 
:::

<br>

### Valid Names 

Object name must consist of letters, digits, and `_`. They cannot begin with a `_` or a digit. 

R will not let hyou use names that are reserved, like `if` or `function`. It is also bad practice to overwrite base functions or other functions that you would like to use (like `mean()`). 

<br> 

# Calling Functions

Speaking of functions, functions in R are objects that reference a block of code that runs when called. 

Functions can have inputs, known as parameters, and can return objects. Neither are mandatory. 

The syntax for calling functions is: 

<br> 

```{r}
#| eval: false

<function_name>(<param1> = "<first_argument>", 
                <param2> - "<second_argument>")
```

<br>

where `function_name` is the name of the function, `param` is the first parameter of the function, and `"first_argument"` is the first user-provided input into the function. 

Note that we are using an `=` sign here; when calling or creating functions, we use `=` to assign the arguments to the parameters. 

<br> 

::: {.callout-tip}
### Using Help

Remember that you can type `?<function_name>` or `help(<function_name>)` into the console to look at the purpose of specific functions and what arguments are valid. 
:::
