---
title: ""
editor: source
---

# Exercise 1

All exercises live in the `exercises/` folder. Navigate to that folder in the `Files` pane and click on `01_exercise.R`. 

## Exercise 1.1

::: panel-tabset

### Problem

- Create the following vectors: 

  - `x`: with values `1`, `3`, `5`, `4`, `5`\
  - `y`: with values `2`, `3`, `1`, `1`, `6` 
  
- Calculate the length of x. Calculate the length of y.  

- Calculate the sum of `x` and `y` such that each element of `x` is added to the corresponding element of `y`. 

- Print the second elemend of `x` using an index.  

- Print that same element using a comparison operator to subset instead. *Hint:* use the value in the second element.  

- Print all elements of `y` less than `3`.  

- Print all elements of `x` less than `5` and greater than `1`.

### Solution

- Create the following vectors: 

  - `x`: with values `1`, `3`, `5`, `4`, `5`\
  - `y`: with values `2`, `3`, `1`, `1`, `6` 
  
  <br>
  
```{r}
# create x and y

x <- c(1, 
       3, 
       5, 
       4, 
       5)

y <- c(2, 
       3, 
       1, 
       1, 
       6)
```
  
<br>
  
- Calculate the length of x. Calculate the length of y.  

```{r}
# print length of x
length(x)

# print length of y
length(y)
```

<br>
- Calculate the sum of `x` and `y` such that each element of `x` is added to the corresponding element of `y`. 

```{r}
# calculate sum of x and y, vectorized
x + y
```
<br> 

- Print the second elemend of `x` using an index.  

<br>

```{r}
# print second element
x[2]
```

<br>

- Print that same element using a comparison operator to subset instead. *Hint:* use the value in the second element.  

```{r}
# print if x is 3
x[x==3]
```
<br>

- Print all elements of `y` less than `3`.   

```{r}
# print if x is less than 3
x[x<3]
```

<br>


- Print all elements of `x` less than `5` and greater than `1`.

```{r}
# print if x is greater than 5 and less than 1
x[x < 5 & x > 1]
```
:::

## Exercise 1.2 - Challenge

::: panel-tabset

### Problem 

- Create a `numeric` vector `z` with elements `1`, `3`, `NA`, `4`, `6`, `NA`.

- Remove the missing values within `z` and print the result. 

<br> 

### Solution

- Create a `numeric` vector `z` with elements `1`, `3`, `NA`, `4`, `6`, `NA`.
<br> 
```{r}
# create z

z <- c(1, 
       3, 
       NA,
       4,
       6,
       NA)
```

<br> 

- Remove the missing values within `z` and print the result. 

```{r}

# remove missing values

z<- z[!is.na(z)]

z
```

<br>
