---
title: ""
editor: visual
---

```{r}
#| echo: false
#| message: false 
#| warning: false

library(tidyverse)
```

# Exercise 2

All exercises live in the `exercises/` folder. Navigate to that folder in the `Files` pane and click on `02_exercise.R`. 

In this exercise we'll continue to use the `starwars` tibble from the `dplyr` package. 

## Exercise 2.1

### panel-tabset

### Problem

#### **Part 1**: Sort the `starwars` dataset, descending by `height`. Assign the new dataset to `starwars_1`. 
#### **Part 2**: Based on the values of `height`, `filter()` `starwars_1` such that it only contains 10 observations. Assign the new dataset to `starwars_2`.

#### **Part 3**: Get the average height in the `starwars_2` tibble. 

<br>

::: {.callout-note}
You may need to set `na.rm = TRUE` in the `mean()` function. 
:::

<br>

### Solution

#### **Part 1**: Sort the `starwars` dataset, descending, by `height`. Assign the new dataset to `starwars_1`. 

```{r}
starwars_1 <- starwars |> 
  arrange(desc(height))
```

#### **Part 2**: Based on the values of `height`, `filter()` `starwars_1` such that it only contains 10 observations. Assign the new dataset to `starwars_2`. 

```{r}
starwars_2 <- starwars_1 |> 
  filter(height >= 202)
```


#### **Part 3**: Get the average height in the `starwars_2` tibble. 

```{r}
starwars_2 |> 
  summarise(mean_height = mean(height, na.rm = TRUE))
```

<br>

:::

