---
title: ""
editor: source
---

```{r}
#| echo: false
#| message: false
#| warning: false

library(tidyverse)
```

# Exercise 9

All exercises live in the `exercises/` folder. Navigate to that folder in the `Files` pane and click on `09_exercise.qmd`. 

## Exercise 9.1  

::: panel-tabset

### Problem

- Install the `gapminder` package and load it. 

- Filter the `gapminder` data frame for rows that have a country of `"United States"`. Assign it to something other than `gapminder`. 

- Graph the population of the United States over time using both points and lines. 

- Increase the size of the points on the graph. Change the shape of the points to `"diamond filled"`. Make the lines blue, the line of the points red, and the inside of the points green. Set the theme to `theme_minimal()`. 

### Solution 

- Install the `gapminder` package and load it. 

<br>

```{r}
# install.packages("gapminder)  
library(gapminder)
```

<br>  

- Filter the `gapminder` data frame for rows that have a country of `"United States"`. Assign it to something other than `gapminder`. 

<br>  

```{r}
us <- gapminder |> 
  filter(country == "United States")
```

<br>  

- Graph the population of the United States over time using both points and lines. 

<br> 

```{r}
ggplot(data = us, mapping = aes(year, pop)) + 
  geom_point() + 
  geom_line()
```

<br> 

- Increase the size of the points on the graph. Change the shape of the points to `"diamond filled"`. Make the lines blue, the line of the points red, and the inside of the points green. Set the theme to `theme_minimal()`. 

```{r}
ggplot(data = us, 
       mapping = aes(year, pop)) + 
  geom_point(size = 1, 
             shape = "diamond filled", 
             color = "red", 
             fill = "green") + 
  geom_line(color = "blue") + 
  theme_minimal()
```

:::

## Exercize 9.2 - Challenge

::: panel-tabset

### Problem

- Starting with the `gapminder` data frame, visualize countries in the Americas' life expectancy over time on the same graph, using the `color` aesthetic. 

- Do the same thing, except do not use color to designate individual countries. Find a different way to separate out the groups. 

### Solution

- Starting with the `gapminder` data frame, visuaolize countries in the Americas' life expectancy over time on the same graph, using the `color` aesthetic. 

<br>  

```{r}
gapminder |> 
  filter(continent == "Americas") |> 
  ggplot(mapping = aes(x = year, y = lifeExp, color = country)) + 
  geom_point() + 
  geom_line()
```

<br> 

- Do the same thing, except do not use color to designate individaul countries. Find a different way to sparate out the groups. 

```{r}
gapminder |> 
  filter(continent == "Americas") |> 
  ggplot(mapping = aes(x = year, y = lifeExp, group = country)) + 
  geom_point() + 
  geom_line()
```

::: 


