---
title: ""
editor: source
---

```{r}
#| echo: false
#| message: false
#| warning: false

library(tidyverse)
```

# Scales  

Scales are an important feature when creating graphics. With them, we can modify our `x` and `y` axes, set limits, change the colors we want to use, and create better labels. 

## `scales_<axis>_continuous()` 

`scale_x_continuous()` and `scale_y_continuous()` modify the `x` and `y` axes respectively when we are using variables. 

They take in the same parameters. 

Some of the most used parameters are: 

- `breaks` 
- `limits` 
- `labels` 
- `expand` 

`ggplot()` will create default scales created from the data. 

```{r}
ggplot(data = diamonds, mapping = aes(carat, price)) + 
  geom_point()
```

Let's see how each parameter might change how the graph looks. 

### `breaks` 

`breaks` assigns where tic marks go on scale. Most often we'll use a numeric vector to specify breaks. 

```{r}
ggplot(data = diamonds, mapping = aes(carat, price)) + 
  geom_point() + 
  scale_y_continuous(breaks = seq(0, 20000, 2000))
```

### `limits` 

`limits` cuts off the scale at specified points. Most often we'll use `c(<from>, <to>)` syntax to modify limits. 

```{r}
ggplot(data = diamonds, mapping = aes(carat, price)) + 
  geom_point() + 
  scale_y_continous(breaks = seq(0, 20000, 2000), 
                    limits = c(0, 10000))
```  

### `labels` 

`labels` modifies text for the tic marks in the scale. Most often we'll use `scales::percent` 


